Open-Source machine learning BANTER acoustic classification of beaked whale echolocation pulses

### 2023-08-31 manuscript/Supplement.qmd
update draft manuscript for submission to RPTS
update draft manuscript and RPTS submission form. Modify Hawaii figures to show top 10 heat map.

---
title: Supplement for Open-Source machine learning BANTER acoustic classification of beaked whale echolocation pulses
author: Shannon Rankin
date: last-modified
output: pdf
---

>>>>

---
title: Supplement for Open-Source machine learning BANTER acoustic classification of beaked whale echolocation pulses
author: Shannon Rankin
date: last-modified
output: pdf
#format: docx
---

### 2023-08-31 manuscript/manuscript.qmd
update draft manuscript for submission to RPTS
update draft manuscript and RPTS submission form. Modify Hawaii figures to show top 10 heat map.

For the SAtlantic dataset, the most accurate model considered the echolocation pulse metrics and the IPI for an event (EC_IPI Model, Detector Model: *sampsize* = 4, *ntree* = 10,000; Event Model: *sampsize* = 1, *ntree* = 100,000) and provided an overall correct classification rate of 88.7% for all five species. Classification scores ranged from a low of 50% (Sowerby's beaked whale) to a high of 100% (True's beaked whale, @fig-satlantic); all classification results were greater than expected (Priors in @fig-natlantic a). The sample sizes for the BANTER model were kept low (event sampsize = 1) to retain Sowerby's beaked whale, which had a sample size of 2. Analysis with a larger sample size resulted in improved classification results for Cuvier's and Gervais' beaked whale (see Supplement Fig. 3 for EC_IPI_ALT), but resulted in the loss of Sowerby's in the final model. Results from the EC_IPI model are presented here; results from the EC and EC_IPI_ALT model can be found in supplementary materials (Supplement Fig. 2, 3).

>>>>

For the SAtlantic dataset, the most accurate model considered the echolocation pulse metrics and the IPI for an event (EC_IPI Model, Detector Model: *sampsize* = 4, *ntree* = 10,000; Event Model: *sampsize* = 1, *ntree* = 100,000) and provided an overall correct classification rate of 88.7% for all five species. Classification scores ranged from a low of 50% (Sowerby's beaked whale) to a high of 100% (True's beaked whale, @fig-natlantic); all classification results were greater than expected (Priors in @fig-natlantic a). The sample sizes for the BANTER model were kept low (event sampsize = 1) to retain Sowerby's beaked whale, which had a sample size of 2. Analysis with a larger sample size resulted in improved classification results for Cuvier's and Gervais' beaked whale (see Supplement Fig. 3 for EC_IPI_ALT), but resulted in the loss of Sowerby's in the final model. Results from the EC_IPI model are presented here; results from the EC and EC_IPI_ALT model can be found in supplementary materials (Supplement Fig. 2, 3).
